#Erik Turchi Lab 1
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#



#Q1 import csv files
```{r}
IBM <- read.csv("IBMStock.csv", header = TRUE)
GE <- read.csv("GEStock.csv", header = TRUE)
ProcterGamble <- read.csv("ProcterGambleStock.csv", header = TRUE)
CocaCola <- read.csv("CocaColaStock.csv", header = TRUE)
Boeing <- read.csv("BoeingStock.csv", header = TRUE)
```



#Q2 check each data field using str() function
```{r}
str(IBM)
str(GE)
str(ProcterGamble)
str(CocaCola)
str(Boeing)
```



#Q3 Clean Data to as.Date Function
```{r}
GE$Date = as.Date(GE$Date, "%m/%d/%y") 
IBM$Date = as.Date(IBM$Date, "%m/%d/%y") 
ProcterGamble$Date = as.Date(ProcterGamble$Date, "%m/%d/%y") 
CocaCola$Date = as.Date(CocaCola$Date, "%m/%d/%y") 
Boeing$Date = as.Date(Boeing$Date, "%m/%d/%y") 
```

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
#Warmup and Basic Q's



#1. How many rows of data are in each dataset?
```{r}
str(Boeing)
str(IBM)
str(GE)
str(ProcterGamble)
str(CocaCola)
```
#A1:  using the str() function it shows in?the console 480 obj(or in this case rows) per data set, as well as 2 columns



#2. What is the earliest/latest year in our datasets?
```{r}
head(Boeing, 1)
tail(Boeing, 1)

head(CocaCola, 1)
tail(CocaCola, 1)

head(ProcterGamble, 1)
tail(ProcterGamble, 1)

head(IBM, 1)
tail(IBM, 1)

head(GE, 1)
tail(GE, 1)
```
#A2:  using the head() and tail() funtions we can assign a value o? 1 column to show the earliest and latest peice of data in the datasets
#     other datasets using mismatched dates and times can be filtered using queries such as dateORder = order(date) and than arrange(dateOrder)



#3. For the period above what is the average stock price of Coca Cola?
```{r}
library(ggplot2)
library(gplots)
plotmeans(StockPrice ~ Date, data = CocaCola)
```
#we can display the average or 'mean' stockprice over the ent?re dataset of 'CocaCola' with plotmeans 

```{r}
mean(CocaCola[["StockPrice"]])
```
#A3:  The mean of the StockPrice Column in CocaCola is 60.03



#4. What is the maximum price of IBM during this period?
```{r}
max(IBM["StockPrice"])
```
#A4:  The max stock price of IBM is 438.90



#5. What is the standard deviation of P&G stock price over this period?
```{r}
sd(ProcterGamble[["StockPrice"]])
```
#A5: ProcterGamble's standard variation (sq root of varience) is displayed with sd() function and is 18.19



#6. What is the median price of Boeing in the last 5 years for which we have data?
```{r}
Boeing2 <- read.csv ("Boeing2004Up.csv", header = TRUE)
Boeing2$Date = as.Date(Boeing2$Date, "%m/%d/%y") 
median(Boeing2[["StockPrice"]])
```
#A6: The median of Boeing from the past 5 years of collected data as 64.96



#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
#Basic Plotting Questions



#Part 1:  Plot the StockPrice of Coca-Col? on the Y-axis across Date on the x-axis using the basic plot() 
```{r}
plot(StockPrice ~ Date, data = CocaCola, frame = FALSE)
```
#Q1 :  Identify the year during which Coca-Cola had the highest/lowest stock price?
#A1 :  The highest Stock Price Year for coca cola was 1973 and its lowest year being 1980

#Q2 :  What calendar year did it look to have the biggest (Year-over-Year) percentage increase?
#A2 :  The highest percentage increase Year OVer Year was 1971 to 1972 with an estimated 55% growth span



#Part 2 : Adding P&G
```{r}
plot(StockPrice ~ Date, data = CocaCola, type = 'l', frame = TRUE, col="red")
lines(ProcterGamble$Date, ProcterGamble$StockPrice,col = "blue")
```
#Q1 : In March of 2000 the stock market plummeted as the tech bubble burst. Using the plot above,
#     which company’s stock dropped more (relatively – i.e. percentage-wise)?
#A1 :  P&G's Stock Dropped More percentage wise than CocaColas

#Q2 :  In the year 1983 which company stock was going up? Which was going down?
#A2 :  In 1983 CocaColas Stocks Were on the rise as P&G's took a sharp drop

#Q3 :  Across the entire time period shown in your plot which stock had a generally lower price?
#A3 :  Across the entire plot CocaColas Stocks were a generally lower price



#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#

#Data Visualization from 1995 - 2005:
#~~~~#
#Instead of looking at the plot across the entire date range, we want to see what’s happening between
#1995-2005. Remember, you can use the matrix notation [rows, columns] to subset data.
#~~~~#
```{r}
plot(CocaCola$Date[301:432],CocaCola$StockPrice[301:432],type = 'l', col = 'red', ylim=c(0,210))
lines(IBM$Date[301:432],IBM$StockPrice[301:432], col = 'Blue')
lines(ProcterGamble$Date[301:432],ProcterGamble$StockPrice[301:432], col = 'green')
lines(Boeing$Date[301:432],Boeing$StockPrice[301:432], col = 'orange')
lines(GE$Date[301:432],GE$StockPrice[301:432], col = 'purple')
```

#Q1 :  Which stock price fell the most right after the tech bubble of March 2000?
#A1 :  GE's Stock Fell the most right after the 2000 bubble


#Q2 :  What stock had the highest maximum price between 1995-2005?
#A2 :  IBM had the highest price maximum between 1995 and 2005


#Q3 :  Afewyearsbeforethetechbubbleof1997,therewasanotherstockmarketcrashtriggedby
#      economic crisis in Asia in October of 1997. If you compare stock prices from September 1997 to
#      November 1997, which companies saw a decrease in price? Which company experienced the
#      biggest decrease?
#A3 :  The companies experiencing a decrease were: IBM, P&G, Boeing, and GE
#      the largest drop experienced was P&G
#      Coca Cola was the only comany that didnt see a massive drop just a tiny bump is a steady growth period


#Q4 :  Whichstock seemedtoprovidethebestreturn(i.e.increaseinprice)between2004-2005?
#A4 :  The stock providing the best return in this era belonged to Boeing


#Q5 :  Between 1995-2005, which company had the biggest delta between the maximum and minimum stock price?
#A5 :  IBM had the largest Delta between the max and Min of the stock price


#Q6 : Which two companies’ stock price seem to be the most correlated (i.e. move up/down together)?
#A6 :  the most coorilated stocks would be GE and P&G


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#

#Monthly Trend Analysis

```{r}
tapply(IBM$StockPrice, months(IBM$Date), mean, simplify = TRUE)
```
```{r}
tapply(ProcterGamble$StockPrice, months(ProcterGamble$Date), mean)
```
```{r}
tapply(Boeing$StockPrice, months(Boeing$Date), mean)
```
```{r}
tapply(GE$StockPrice, months(GE$Date), mean)
```

#Q1 :  For IBM ,compare the average stock price for each month tothe its overall average stockprice
#      and identify all the months for which IBM historically had a higher stockprice(we call this over indexing)? Which month over-indexed the most?
#A1 :  IBM's Historic 'good' months are Janu?ry, February, March, April, and May, a historically successful first half to the year


#Q2 :  Repeatthetapply()functionyouusedtosolve thelastquestionforeachofthe4remaining
#      companies. Do any of two or more companies have their highest stock price in the same months
#      as each other? Which companies and months does this happen for?
#A2 :  G&E, coca cola, and IBM have very similar high points in the same months: January, February, March, April, and May respectivly


#Q3 :  What trend do you see for the months of December vs January for each company? Is there an
#      over-arching trend that applies to all companies when comparing all historical December vs
#      January stock prices?
#A3 :  December for every company ?s a low point aside from but not excluding coca cola(3rd lowest month) It seems december to january 
#      is always a jump upwards in stock prices with most companies almost having their highest averaging months during january


