```{r}
housing <- read.csv("housing.csv")
library(ggplot2)
library(dplyr)
```

#Time to Check the data
```{r}
head(housing)
```

#checking data types
```{r}
str(housing)
```

#check statistics
```{r}
summary(housing)
```

#we look into which data has NULL values
```{r}
colSums(is.na(housing))
```

#for all data na in total_bedroooms replace with median of total_bedrooms as outliers wont come into affect
```{r}
housing$total_bedrooms[is.na(housing$total_bedrooms)] <- median(housing$total_bedrooms, na.rm = TRUE)
```
```{r}
summary(housing$total_bedrooms)
```

#creating dummy variables for ocean_proximity
```{r}
library(fastDummies)
```
```{r}
h1 <- dummy_cols(housing$ocean_proximity, remove_first_dummy = TRUE)
```

#binding dummies to housing dataset
```{r}
housing <- cbind(housing,h1[-c(1)])
```

# find average bedrooms per household
```{r}
housing$mean_bedrooms = housing$total_bedrooms/housing$households
```
#find average rooms per household
```{r}
housing$mean_rooms = housing$total_rooms/housing$households
```
#create omit for columns total_rooms and total_bedrooms; omit ocean_proximity column now that dummy variables have been created
```{r}
omit = c("total_rooms", "total_bedrooms","ocean_proximity")
```
#use omit to take the total housing and bedrooms out of dataset since we have mean room and bedroom now created
```{r}
housing <- housing[ , !(names(housing) %in% omit)]
```

```{r}
str(housing)
```
#load libraries for data visualization
```{r}
library(ggplot2)
library(gplots)
```
```{r}
library(plotly)
packageVersion('plotly')
```


```{r}
ggplot(housing, aes(housing_median_age)) +
  geom_histogram(bins = 100)
```

```{r}
ggplot(housing, aes(housing_median_age)) +
  geom_density()
```

```{r}
p1 <- plot_ly(housing, x = ~median_income)%>% add_histogram()
p2 <- plot_ly(housing, x = ~median_house_value)%>% add_histogram()
subplot(p1,p2)

#on the left:  median income distribution
#on the right: median house value distribution
```

```{r}
ggplot() +
  geom_point(data = housing, aes(x = median_house_value, y = population), color = "red")
```


```{r}
install.packages(c("ggplot2", "devtools", "dplyr", "stringr"))
```
```{r}
install.packages(c("maps", "mapdata"))
```
```{r}
library(ggplot2)
library(maps)
library(mapdata)
```
```{r}
library(ggplot2)
library(leaflet)
library(leaflet.extras)
```

```{r}
leaflet(housing)%>%
  addWebGLHeatmap(size=35000)%>%
  addProviderTiles(providers$CartoDB.DarkMatter)
```

